Location and Scale-Invariant Power Transformations for Transforming Data to

### 2025-04-02 manuscript_rmarkdown/manuscript.Rmd
Expand description of text.

To determine the weighting function parameters for each of the nine combinations,
$m_d=100$ sequences $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
drawn from randomly parametrised asymmetric generalised normal distributions.
Each distribution was parametrised with a randomly chosen skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
To form each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, 
with $\gamma \sim U\left(2, 4\right)$, resulting in a set of sequences with between $100$ and $10000$ elements.

@@

## Simulated data

[Generated data: clean, n = 30, n = 100, n = 1000, lambda = -1.00, 1.00, 3.00]

@@

\begin{table}
\begin{center}
\caption{
Comparison of average rank between transformation methods based on either residual error (without outliers) or residual error of the central portion
(with outliers; $\kappa = 0.70$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
}
\label{tab:comparison_methods_simulations}
\begin{tabular}{l | l r r r r r r}

>>>>

To determine the weighting function parameters for each of the nine combinations,
$m_d=100$ sequences $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
drawn from randomly parametrised asymmetric generalised normal distributions.
Each distribution was parametrised with a random skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
To form each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, 
with $\gamma \sim U\left(2, 4\right)$, resulting in a set of sequences with between $100$ and $10000$ elements.

@@

## Assessing transformation using simulated data

Three datasets with 10000 sequences each were created to assess transformation
to normality. For each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$
elements were randomly drawn, with $\gamma \sim U\left(1, 4\right)$, 
resulting in sequences with between $10$ and $10000$ elements.

- A \textit{clean} dataset with each sequence drawn from $\mathcal{N}(0, 1)$ and
  transformed using an inverse power transformation with randomly drawn 
  transformation parameter $\lambda \sim U\left(0.00, 2.00 \right)$. 
  $\left(\phi_{\text{BC}}^\lambda\right)^{-1}$ and $\left(\phi_{\text{YJ}}^\lambda\right)^{-1}$
  were used as inverse transformations for assessing Box-Cox and Yeo-Johnson transformations,
  respectively. Prior to inverse transformation, sequences for Box-Cox transformations
  were shifted into the positive domain, with minimum value $1 \cdot 10^{-4}$.

- A \textit{dirty} dataset with each sequence drawn from $\mathcal{AGN}\left(\mu = 0, \sigma = 1/\sqrt{2}, \alpha, \beta \right)$,
  with randomly drawn skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
  and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.

- A \textit{shifted} dataset with each sequence drawn from $\mathcal{AGN}\left(\mu = 100, \sigma = 10^{-3} \cdot 1/\sqrt{2} , \alpha, \beta \right)$,
  with randomly drawn skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
  and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.

A dataset with outliers was created for each of the above datasets by replacing
10 percent of elements in each sequence, as described earlier.

In addition to no power transformation and location and scale-invariant power transformations,
conventional and Raymaekers and Rousseeuw's robust adaptation [@Raymaekers2024-zf] were assessed.
For the latter two methods, normalisation before standardisation using was additionally assessed using
the following two methods:

1. z-standardisation: $x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma$, with $\mu$ and $\sigma$ the mean and standard deviation of sequence $\mathbf{X}$.

2. robust scaling: $x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)$,
   with $\text{IQR}$ representing the interquartile range of $\mathbf{X}$.

This results in nine types of power transformation. Transformation parameters
were optimised for each sequence. Subsequently the sum of residual errors 
and sum of residual errors of the central portion ($\kappa = 0.70$) were computed
for each sequence after transformation. Then, each method was ranked according 
to the sum of residual errors (data without outliers) or the sum of residual
errors of the central portion (data with outliers). The average rank of each
transformation method was computed over all 10000 sequences in each dataset.
For Yeo-Johnson transformations average ranks are shown in Table \ref{tab:comparison_methods_simulation_yeo_johnson}.

@@

\begin{table}
\begin{center}
\caption{
Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
(with outliers; $\kappa = 0.70$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
}
\label{tab:comparison_methods_simulations_yeo_johnson}
\begin{tabular}{l | l r r r r r r}
